Bayes Optimal Informer Sets for Drug Discovery\thanks{abc}

### 2020-04-09 BOISE/Paper_draft/draft.tex
update draft

...on what to do next

...on connection to Bayesian adaptive design (e.g.
two-stage finite decision problems ; section 12.3.1 in 
Parmigiani and Inoue...connection to sequential clinical trials...

The idea of BOISE comes from Bayes decision theory (Parmigiani and Inoue 2009) and adaptive clinical trial design (Berry 2006). In section 12.3 of Parmigiani and Inoue (2009), a general procedure is formalized to solve two-stage finite decision problems: The second stage decision problem should be solved first, assuming that action and outcome in previous stage are given. Then we can go back to first stage decision problem with the help of optimal action in second stage. If we formalize the multistage decision problem as a decision tree, as Parmigiani and Inoue did in their book, this procedure seems like a recursion from its leaves to its root. In BOISE, informer set selection is the first stage problem and top set selection (or ranking) is the second stage problem. Therefore, we first look for the optimal top set rule $T^*(A,x_A,x_0)$, assuming we know the first stage action $A$ and outcome $x_A$, and then we step back to find best informer set rule with optimal top set rule $T^*(A,x_A,x_0)$.

When we seek the best informer set rule in first stage problem, ideas of adaptive design are edifying. Berry (2006) has formalized an adaptive trial design idea based on Bayesian predictive distribution, and this idea finally develops into I-SPY 1 trial (Gomez {\em et al.} 2007, 2008) and I-SPY 2 trial (Barker {\em et al.} 2009). In these clinical trials, drugs (or therapies) will be sequentially evaluated based on their predictive distribution. Every time when new patients come into study, decisions are made to drop, graduate, or continue testing on certain drugs. Evaluating drugs based on predictive distribution is to calculate the loss of all possible outcomes in the future, then take the average with respect to their conditional probability given previous outcomes. In BOISE, this idea is used to find optimal informer set $A$, where we pretend we have the future test result on inform set $A$, that is $x_A$, and take the integral of future loss ${\rm PEL}_2(x_0,x_A,A)$, with respect to conditional distribution $p(x_A\,|\,x_0)$. An adaptive design idea is proved to be more effective than standard designs in clinical trials, and now is also shown to be promising in informer set selection problem. 
%\singlespacing
%\setlength\bibsep{0pt}
%\bibliographystyle{my-style}
%\bibliography{Placeholder}

\clearpage
\appendix

>>>>

The idea of BOISE comes from Bayes decision theory (Parmigiani and Inoue 2009) and adaptive clinical trial design (Berry 2006). In section 12.3 of Parmigiani and Inoue (2009), a general procedure is formalized to solve two-stage finite decision problems: The second stage decision problem should be considered first, assuming that action and outcome in previous stage are given. Then we can go back to first stage decision problem with the help of optimal action in second stage. If we formalize the multistage decision problem as a decision tree, as Parmigiani and Inoue did in their book, this procedure seems like a recursion from its leaves to its root. In BOISE, informer set selection is the first stage problem and top set selection (or ranking) is the second stage problem. Therefore, we first look for the optimal top set rule $T^*(A,x_A,x_0)$, assuming we know the first stage action $A$ and outcome $x_A$, and then we step back to find best informer set rule with optimal top set rule $T^*(A,x_A,x_0)$.

When we seek the best informer set rule in first stage problem, ideas of adaptive design are edifying. Berry (2006) has formalized an adaptive trial design idea based on Bayesian predictive distribution, and this idea finally develops into I-SPY 2 trials (Barker {\em et al.} 2009). In these clinical trials, drugs (or therapies) will be sequentially evaluated based on their predictive distribution. Every time when new patients come into study, decisions are made to drop, graduate, or continue testing on certain drugs. Evaluating drugs based on predictive distribution is to calculate the loss of all possible outcomes in the future, then take the average with respect to their conditional probability given previous outcomes. In BOISE, this idea is used to find optimal informer set $A$, where we pretend we have the future test result on inform set $A$, that is $x_A$, and take the integral of future loss ${\rm PEL}_2(x_0,x_A,A)$, with respect to conditional distribution $p(x_A\,|\,x_0)$. An adaptive design idea is proved to be more effective than standard designs in clinical trials, and now is also shown to be promising in informer set selection problems.

One important issue in BOISE computation is MCMC sampling from Dirichlet Process. For conjugate models like in (\ref{eq:model}), we can collapse the computation by integrating parameters out and MCMC will mix well. One way to check for mixing properties of the Markov chain in DPMM is to check for the chain over numbers of clusters. Figure \ref{fig:mcmc} shows the plot and estimated auto-correlation of this chain over cluster numbers on PKIS1 and GDSC.

\begin{figure}[!ht]
\caption{\label{fig:mcmc} 
{\bf Mixing properties of MCMC sampling over clusters on PKIS1 and GDSC} The plots are based on the chain over numbers of clusters in each DPMM sample.}
\centering
\includegraphics[width=5.0in]{DPMM_performance.png}
\end{figure}

From the ACF plots in Figure \ref{fig:mcmc}, Markov chains on PKIS1 and GDSC both mix well as there is tiny auto-correlation among consecutive samples. However, from left two plots of cluster numbers in each sample in Figure \ref{fig:mcmc}, Markov chain on PKIS1 seems to be more stable than it on GDSC, and we believe this is caused by the nature of the GDSC data set since the large volatility still remains when we run a longer MCMC or change the prior mass. The volatility also corresponds to the difference between performances of BOISE on PKIS1 and GDSC: Figures  \ref{fig:nef} to \ref{fig:auc2} and Figure \ref{fig:gdsc} show that overall BOISE perform better on PKIS1 data set than on GDSC. This difference between PKIS1 and GDSC seems to be caused by some overwhelming data points that can either be assigned to some existing clusters or form a new cluster. In the future, we may allow for multiple cluster assignments for a single data point instead of forcing it to stay in only one cluster, and there is a corresponding extension of Chinese Restaurant process, named as Indian Buffet process, that admits multiple labels in DPMM. 
%\singlespacing
%\setlength\bibsep{0pt}
%\bibliographystyle{my-style}
%\bibliography{Placeholder}

\appendix
